---
title: "Seaweeds through space and time: building tools for managing the Falkland Islands' unique coastal biodiversity"
author: "Rob Mrowicki"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    fig_caption: yes
    fig_crop: yes
    number_sections: true
  word_document:
    fig_caption: yes
    number_sections: true
header-includes:
  - \usepackage{placeins}  # \FloatBarrier
---

```{r init, include=FALSE}
source('./scripts/00_init.R')  # run initialisation script
```

```{r source_scripts, include=FALSE}
# run data scripts:
source('./scripts/01_import_tidy_manipulate_data.R')
source('./scripts/02_species_data_analysis.R')
source('./scripts/03_sites_data_analysis.R')
source('./scripts/04_plot_figures.R')
```

# Methods

## Data processing

### Seaweed specimen records

Seaweed species records for the Falkland Islands were obtained from (1) historical and recent collections in the NHM algal herbarium, including specimens collected by Mystikou et al., and (2) contemporary specimens collected during the DPLUS068 'Falklands marine forests' project (2018--19).

Georeferencing (i.e. assigning point coordinates and associated 'extent') of historical records, and other records without specific coordinates, followed the _NHM Georeferencing Guidelines_. Contemporary records with GPS coordinates but no measure of GPS error were assigned an extent of 30&nbsp;m (Figure 1).

Records with dubious and/or ambiguous localities (e.g. 'Valparaiso & Falklands') were excluded, as were records for specimens collected as 'drift' (i.e. unattached to substratum).

Where collection date was given as a range (e.g. 'from July to November 1910'), the mean of the start and end date was calculated. Where possible, collection date (or date range) for records lacking date information was inferred based on collector and/or expedition (e.g. from '1839 to 1843') for specimens collected by J.D. Hooker or D. Lyall during the Ross Antarctic Expedition. Records for which collection date could not be inferred were excluded. All remaining records were categorised according to year group as follows: `r year_grps`.

Overall, `r prettyNum(nrow(dd_specimens), ",")` records from `r paste(range(dd_specimens$year), collapse = "--")` were used for subsequent analyses (Figure 2).


### Spatial data

Grid cells for assessing the spatial distribution of taxa (Figure 1) were derived via rasterisation of Falkland Islands coastline vector data (sourced from [OpenStreetMap](https://osmdata.openstreetmap.de/data/coastlines.html) on 10/09/2019) based on a cell size of `r grid_res/10^3` $\times$ `r grid_res/10^3`&nbsp;km. The appropriate grid cell size was determined from the square root of the mean Voronoi polyogon area calculated for all sites (for which 'extent' $\leq$ `r ext_lim/10^3`&nbsp;km) located at least `r nf_dist/10^3`&nbsp;km from the centre of Stanley (`r paste(stanley_coords, collapse = ", ")`), using the `voronoi.polygons` function in the R package `SDraw`.


## Data analysis

### Species-level assessment

From a total of `r length(all_taxa)` taxa across all records, `r length(use_taxa)` were selected for species-level assessment (Table 1). Additionally, `r length(use_grps)` taxonomic groups for which there was some uncertainty over the identities of constituent taxa (_n_ = `r length(unlist(all_grps_taxa[use_grps]))`) were selected for distribution mapping:

```{r grps_taxa_table, echo = TRUE}
# Constituent taxa within selected taxon groups:
all_grps_taxa[use_grps]
```

For each taxon, the extent of occurrence (EOO) and area of occupancy (AOO) were determined for the most recent year group (`r last(year_grps)`). These calculations were only possible for taxa with records from at least three locations. EOO was calculated as the area of the $\alpha$-shape derived using the `alphashape` function in the R package `alphahull`, with $\alpha$ arbitrarily defined as the maximum spatial extent of records (by either latitude or longitude) multiplied by three. AOO based on occupancy of 2 $\times$ 2&nbsp;km grid cells was predicted from a log-linear model of AOO versus occupancy for a range of grid sizes achieved via successive doubling of the base grid size, following _IUCN Red List guidelines_.

The number of specimen records (whereby each 'record' corresponded to a single collection 'event', i.e. unique combination of collector, year and site) were summed for each location (= group of sites; see below) across all year groups, for each taxon. Additionally, the number of locations from which each taxon was collected was calculated for the most recent year group.

```{r loc_grps_table, echo = TRUE}
# Classification of sites by location:
loc_grps_locs[loc_grps]
```


### Site-level assessment

Contemporary records from the DPLUS068 'Falklands marine forests' project (2018--19) were used for site-level assessment, excluding sites for which survey duration was less than 30&nbsp;minutes. Seaweed taxonomic richness was calculated for each of the remaining `r length(use_site_codes)` sites, and categorised (based on sample quartiles) as follows: `r rich_grps`.


# Results

## Species-level assessment

Figures 3 and 4 show example distribution maps for selected taxa and taxon groups. Table 2 shows EOO, AOO and number of locations for selected taxa collected in the Falkland Islands during `r last(year_grps)` and Table 3 shows an example summary of the number of records from different locations across all year groups.


## Site-level assessment

Figure 5 shows seaweed taxonomic richness at sites surveyed during the DPLUS068 project during 2018--19.


\newpage

## Figures

\FloatBarrier

![Distribution of historical and contemporary Falkland Islands seaweed specimen records, for which spatial extent $\leq$ `r ext_lim/10^3`&nbsp;km (i.e. excluding unlocated records from 'East Falkland', 'West Falkland', and the 'Falkland Islands'). Circles represent extent assigned through the georeferencing protocol.](./figures/coords_map.pdf)

![Temporal distribution of Falkland Islands seaweed specimen records included in analyses.](./figures/histogram.pdf)

![Example distribution map for a selected taxon, showing collection sites and corresponding grid cell occupancy. Based on the most recent year group only.](./figures/taxon_maps_eg.pdf)

![Example distribution map for a taxon group within which there is still some taxonomic uncertainty, showing collection sites for individual taxa and grid cell occupancy for the taxon group as a whole. Based on the most recent year group only.](./figures/taxon_group_maps_eg.pdf)

![Seaweed taxonomic richness at sites surveyed during the DPLUS068 'Falklands marine forests' project (2018--19).](./figures/rich_map.pdf)


\newpage

## Tables

\FloatBarrier

```{r taxa_table, echo = FALSE, results = 'asis'}
use_taxa %>%
  kable(
    col.names = "Species",
    caption = "Individual taxa selected for species-level assessment."
  )
```

```{r taxa_eoo_aoo_locs_table, echo = FALSE, results = 'asis'}
taxa_eoo_aoo_locs %>%
  # format numerical columns:
  mutate_at(2:3, ~ formatC(., big.mark = ",", digits = 0, format = "f")) %>%
  kable(
    align = "lrrr",  # right align numerical columns
    caption = "Extent of occurrence (EOO), area of occupancy (AOO) and number of locations for selected seaweed taxa collected from the Falkland Islands during the most recent year group only."
  )
```

```{r taxa_st_table, echo = FALSE, results = 'asis'}
taxa_st[["Cladodonta lyallii"]] %>%  # must index list properly to result in data frame
  rownames_to_column("location") %>%  # NB -- having rownames messes up formatting
  kable(
    caption = "Summary of the number of records from different locations across all year groups for the species \\emph{Cladodonta lyallii}."
  )
```


\newpage

## Session info

```{r session_info, include = TRUE, echo = TRUE}
session_info()
```
