---
title: "Seaweeds through space and time: building tools for managing the Falkland Islands' unique coastal biodiversity"
author: "Rob Mrowicki"
date: "`r format(Sys.time(), "%d %B %Y")`"
output:
  pdf_document:
    fig_caption: yes
    fig_crop: yes
---

```{r init, include = FALSE}
source('./scripts/00_init.R')  # run initialisation script
```

```{r source_scripts, include = FALSE}
# run data scripts:
source('./scripts/01_import_tidy_manipulate_data.R')
source('./scripts/02_plot_figures.R')
```

## Data analysis

Georeferencing (i.e. assigning point coordinates and associated 'extent') of historical records, and other records without specific coordinates, followed the _NHM Georeferencing Guidelines_. Contemporary records with GPS coordinates were assigned an extent of 30 m.

Records with dubious and/or ambiguous localities (e.g. 'Valparaiso and Falklands') were excluded.

Where collection date was given as a range (e.g. 'from July to November 1910'), the mean of the start and end date was calculated. Where possible, collection date (or date range) for records lacking date information was inferred based on collector and/or expedition (e.g. from '1839 to 1843') for specimens collected by J.D. Hooker or D. Lyall during the Ross Antarctic Expedition. Records for which collection date could not be inferred were excluded.

Remaining records were categorised according to year group as follows: `r year_grps`.

Records for contemporary specimens collected as 'drift' (i.e. unattached to substratum) were excluded.

An appropriate grid size (`r grid_res/10^3` km) for assessing the spatial distribution of taxa was based on the square root of the mean Voronoi polyogon area calculated for all sites (for which 'extent' $\leq$ `r ext_lim` m) located at least `r nf_dist/10^3` km from the centre of Stanley.

For each taxon, the extent of occurrence (EOO) and area of occupancy (AOO) were determined for the most recent year group (`r year_grps[length(year_grps)]`) in cases where there were records from at least three locations. EOO was calculated as the area of the $\alpha$-shape derived using the `alphashape` function in the R package `alphahull`, with $\alpha$ arbitrarily defined as the spatial extent of records multiple by three. AOO based on occupancy of 2 km $\times$ 2 km grid cells was predicted from a log-linear model of AOO versus occupancy for a range of grid sizes achieved via successive doubling of the base grid size, following _IUCN Red List guidelines_.

The number of specimen records (whereby each 'record' corresponded to a collection 'event', i.e. unique combination of collector, year and site) were summed for each location (= group of sites) across all year groups, for each taxon.


## Figures

![Temporal distribution of Falkland Islands seaweed specimen records included in this study.](./figures/histogram.pdf)

![Distribution of historical and contemporary Falkland Islands seaweed specimen records, for which spatial extent $\leq$ `r ext_lim/10^3` km (i.e. excluding unlocated records from 'East Falkland', 'West Falkland', and the 'Falkland Islands'. Circles represent extent assigned through the georeferencing protocol.)](./figures/coords_map.pdf)
